Manuscript: Detecting differences in Size Spectra

### 2023-02-23 Manuscript.Rmd
MS draft 23 feb 2023

---
title: "Manuscript: Detecting differences in Size Spectra"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: 
  - Justin Pomeranz:
      email: jfpomeranz@gmail.com
      institute: cmu
      correspondence: true
  - name: James R. Junker
    institute: [glrc, lumcon]
  - name: Vojsava Gjoni
    institute: usd
  - name: Jeff S. Wesner
    institute: usd
institute:
  - cmu: Colorado Mesa University, Grand Junction, CO, USA
  - glrc: Great Lakes Research Center, Michigan Technological University, Houghton, MI USA
  - lumcon: Louisiana Universities Marine Consortium, Chauvin, LA USA
  - usd: Dept. of Biology, University of South Dakota, Vermillion, SD, USA
output:
   word_document:
      reference_docx: working_docx_template.docx
      toc: no
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
   pdf_document:
     toc: yes
     keep_tex: true
     pandoc_args:
     - '--lua-filter=scholarly-metadata.lua'
     - '--lua-filter=author-info-blocks.lua'
header-includes:
  - \usepackage{lineno}
  - \usepackage{amsmath}
  - \usepackage{indentfirst}
  - \linenumbers
  - \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
indent: true
linestretch: 1
bibliography: detecting-size-spectra-difference.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

@@

**NOTE** I'm still working on integrating with Zotero and importing all the necessary references (thanks Jim for the help and instructions!). Don't worry about the **??** references in the rendered version for now.

@@

Body size distributions are a fundamental characteristic of communities. In general, size-abundance relationships are expected to scale with the value of -0.75 across all biological communities [@Damuth1981, @Damuth1991, @Damuth1998]. This is thought to be a consequence of simple size-dependent metabolic constraints on organisms’ energy use predicted by the metabolic theory of ecology [@Nee1991, @Brown2004]. The remarkable consistency of these relationships across spatiotemporal scales and ecosystems has led them to be recommended as a "universal" indicator of ecological status [@Petchey2010]. Variation in size-abundance relationships have been documented through space [@pomeranz2022], time [@evans2022], seasonality [@mcgarvey2018], and in response to human activities [@jennings2004]. Likewise, variation in size spectra relationships have been used to explain fundamental differences in how communities are organized. For example, external resource subsidies "bend the rules" and allow higher abundances of large body sizes than would be expected [@perkins2018]. However, recent research has shown that these results may be an artifact of how the data were treated. Edwards et al. (2020) reanalyzed a time series of marine fisheries data and found that previously reported changes through time were actually dependent on the methodology used.

Individual size distributions (ISD *sensu* [@White2007], also referred to as abundance size spectra) are one of the size-abundance relationships commonly used. ISDs represent frequency distributions of individual body sizes within a community, regardless of taxonomy. Generally, there is a negative relationship between individual body size (M) on the *x*-axis  and abundance (N) on the *y*-axis. Theoretical and empirical data support this relationship being described as a simple power law with exponent $\lambda$ in the form of $N \sim M^{\lambda}$, where $\lambda = 2$ [@Sheldon1972, @Andersen2006a]. Commonly, $N$ is the count of body sizes grouped into bins, and $\lambda$ is estimated as the slope from OLS regressions in log-log space of $N$ and the mid point of the body size bin $M_{bin}$: $\log(N_{count}) = \lambda \log(M_{bin})$. Myriad binning methods have been proposed, including linear and logarithmic bin widths. Likewise, some methods rely on the absolute counts in the bins and others employ normalization techniques such as dividing the count by the bin width (especially common with logarithmic binning). As an alternative to binning methods, $\lambda$ can be estimated directly on un-binned data using maximum likelihood techniques.

Previous work has shown that the estimates of $\lambda$ differ between MLE and size-binned OLS techniques. Size-binned OLS methods are particularly sensitive to decisions made in the binning process. Simulation studies have shown that MLE offers consistently more accurate estimates of $\lambda$ (White et al. 2008, Edwards et al. 2017), and reanalysis of empirical data also indicates that the conclusions are dependent on the methodology used [@white2008 @edwards2020]. However, recent empirical analysis of stream macroinvertebrate communities across the National Ecological Observatory Network (NEON, USA) showed that while the estimates of $\lambda$ varied based on method, the relative change across the environmental gradient was consistent across methods [@pomeranz2022]. While there is a growing consensus that MLE methods offer more reliable estimates of $\lambda$, and binning methods result in biased estimates, it remains unclear if these biases are consistent and systematic or stochastic, and whether or not the relative change in ISD parameters is consistent across space and time. In other words, if the data within a study are all treated the same, does a relative change of size-binned OLS slope parameters of 0.1 coincide with a relative change of MLE $\lambda$ estimates of 0.1?

@@

For the first binning method, we created 6 equal logarithmic bins covering the range of body sizes. This method has been used extensively in previous studies. For example see [@perkins2018, @martinez2016, @dossena2012]. In the present study, the count in each bin was normalized by dividing by the bin width to account for the unequal bin sizes. The process of normalizing shifts the estimate by -1. In other words, an un-normalized estimate of -0.75 would result in an estimate of -1.75 when normalizing the data [@sprules2015, @edwards2017, @pomeranz2022]. Throughout the manuscript, the normalized equal logarithmic binning method will be referred to as ELBn.

### Log~2~ Bins: NAS
The second method was similar to ELBn, but bins of log~2~ widths are used. The number of bins is dependent on the range of body sizes present in the data. When working with empirical data with different size ranges this can alter the number of bins per site, and it is recommended to construct the bins based on the global min and max body size present. However, since the data here were simulated from a known size range, the number of bins for each site is identical. The count in each bin is normalized in the same way as described above. This method has been  used extensively in the literature [\e.g., @jennings2004, @Jennings2002, @sprules2015, @pomeranz2019; @mcgarvey2018) and is referred to as the Normalized Abundance Spectrum (NAS).

@@

There was considerable variation in the $\lambda$ estimate across methodologies (Figure 1). The distribution of estimates from the MLE method was almost always symmetrical and centered at the known value of $\lambda$ (Figure 1, blue distribution and dashed line). The distribution of estimates from the binning methods were generally wider, and sometimes showed assymetrical distributions.

@@

@@

The distribution of relationship estimates from the MLE method (Figure 3, blue) is approximately centered at the known relationship value in each of the "windows". The distribution for the MLE estimate is widest in the "steep" window, but is still markedly narrower than the binning methods. In addition to having much wider distributions, the binning methods systematically overestimate the known relationship.

@@

Finally, we were interested in finding if the three methods performed differently with varying values of the known relationship. Here, we present estimates from each method when the known relationship ($\beta_{env}$) across the hypothetical gradient was -0.5, -0.25, and 0, respectively (Figure 4).

![Figure 4. Individual regressions for each rep (N = 1000) and method (columns) for three different known relationship values (rows): 0, 0.25, and 0.5](figures/vary_beta_plot.png)

Likewise, we also present the distribution of the $\beta_{env}$ relationship estimates (Figure 5). All the methods perform similarly and reasonably well when no relationship is present. The performance of the binning methods declines (center of distributions), with relationships being systematically over estimated. Likewise, the width of the distributions increases with stronger relationships across a hypothetical gradient. However, the distribution of estimates from the MLE method is nearly always centered at the known value, and has relatively narrow distributions.

@@

For both empirical data sets, the direction and magnitude of change (i.e. $\beta_1$ coefficients) are generally in agreement. Size spectra parameters consistently increase (become flatter) in the AMD data (Fig. 6A). However, the magnitude of the change across the AMD gradient is weakest using the MLE method, and the width of the estimates ($\pm$ 1 SD) is larger in the binning methods, particularly the ELBn method (Fig. 6B). Likewise, the size spectra parameters consistently increase (become steeper) with increasing temperature across the NEON sites (Fig. 6C). The magnitude of the relationships is weakest in the MLE method, although the uncertainty around the estimates are similar (Fig. 6D).

@@

The $\beta_1$ coefficient estimates for the AMD data are all positive and of relatively similar magnitudes across methods. The MLE estimate has the smallest magnitude and SD (0.063 $\pm$ 0.011 SD), The ELBn method is slightly larger and has a wider SD : (0.072 $\pm$ 0.018). Finally, the NAS method has the largest estimate and an intermediate standard deviation (0.076 $\pm$ 0.013). The range of the gradient in the AMD data is relatively small (9.5), but the absolute change in the size spectra parameter ($\lambda$) across the AMD gradient vary by ~0.13 units depending on the method used (range of absolute change: 0.59 to 0.72).

Similar to the AMD results, the $\beta_1$ coefficient estimates for the NEON data are all negative and have similar magnitudes across methods. The MLE estimate is once again the smallest (-0.006 $\pm$ 0.001 SD). The NAS and ELBn methods are larger at -0.004 $\pm$ 0.002, and -0.002 $\pm$ 0.002, respectively. Although the estimates are similar and have a small magnitude, the range of the gradient in this data is larger (\~30$\circ$C), meaning that the absolute change in $\lambda$ estimates varies by 0.11 units depending on the methods used (range of absolute change across gradient: 0.06 to 0.17).

@@

The relationship between body size and abundance has been extensively studied in a wide range of taxa inhabiting both terrestrial and aquatic ecosystems (reviewed by [@brown1995, GastonBlackburn2000, @White2007]). Empirical patterns are remarkably consistent, and can be explained by the metabolic theory of Ecology [@Brown2004].  Measuring parameters describing the decline in abundance with increasing body size in communities is being done with increasing frequency across ecology. Previous work has investigated the accuracy and inherent biases associated with different estimation methods. However, how these inaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detect variation in size spectra parameters across environmental gradients with confidence. Here, we sampled body sizes from known distributions with varying parameters ($\lambda$) and estimated the coefficient of the relationship ($\beta_{env}$) across a hypothetical gradient in order to assess how the results would vary depending on the method used. Likewise, we compared how the interpretation of previously published results could change depending on the methodology used to estimate the relationship.

Generally, the estimate of the size spectra parameter from the MLE method was always closer to the known $\lambda$ value, regardless of how steep or shallow the $\lambda$ exponent was. This is in agreement with Edwards et at. (2017) which found that the MLE estimate more consistently recovered known $\lambda$ values ranging from -2.5 to -0.5. We further this work by discovering that the magnitude of the error associated with the two binning methods increases as the $\lambda$ value moves away from $\sim$ -1.

It is unsurprising that all methods performed less well in estimating steep values of $\lambda$ (i.e., -2.5). The rapid decline of large body sizes makes it unlikely to sample any individuals with a large body size. Furthermore, the binning methods inherently "lose" data by aggregating individual observations into a single bin.

Likewise, the MLE method also resulted in the most accurate $\beta_{env}$ estimates for the hypothetical relationship in simulated data regardless of the $\lambda$ "window" investigated. Furthermore, the variation around the estimate was always smallest with the MLE method. However, the variation in the MLE estimate did increase with steeper $\lambda$ "windows". This is not surprising, however because steep relationships like this make it very unlikely to sample the largest body sizes, and capturing the "true" relationships without the full range of body sizes present is more difficult (Edwards et al. 2017). Despite this, the MLE method is nearly always the best option to use, likely because all the data are used in the analysis.

Binning methods are easy to use and interpret, which most likely accounts for their wide use in ecological studies. However, aggregating individuals into logarithmic bins removes a large source of the variation within the data by collapsing body size variation into a single value within each bin. For example, all individuals placed into a bin that ranges from 2-4 grams of mass are all treated as having a mass of 3 grams, the midpoint of that bin. Likewise, a single abundance value is taken for this bin, despite that fact that there is almost certainly variation in the abundance of individuals that weigh \~2, \~3, or \~4 grams. By homogenizing the data in this way, it seems that binning methods will always produce noisier results than MLE from the binning process alone, since binning is akin to deleting information that the model could otherwise use. One of the benefits of using MLE is that all data points are retained for use within the model.

The variation in estimated relationships across the empirical datasets varied by ~0.1 units depending on the method used. At first glance this may not appear to be a concerning magnitude of variation. However, when one considers the range of variation reported in many seminal works on variation in size spectra relationships is ~0.1-0.2 (SI Table *xx*), it becomes clear how important an error in estimates of this magnitude is.

Despite the differences in magnitude of the empirical relationship coefficients, they were generally in the correct direction and of a similar magnitude. This suggests that previously reported significant changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible. Given that all of the data within a study is treated identically, the over all change in size spectra parameters is likely reasonable. However, the biases and inconsistencies in relationship estimates presented here suggest that it would be difficult if not impossible to directly compare the relative changes across different published studies which use different methods. However, the relatively systematic error of $\lambda$ estimates across known exponents suggests that it may be possible to correct previous estimates onto the same scale (Figure 2).

Despite this possibility, publication of individual body size data with future studies of size spectra relationships would greatly aid in our ability to generalize changes to this fundamental aspect of community organization across spatiotemporal scales and in response to environmental conditions.

@@

The MLE method outperformed both of the binning methods under nearly any measure. With the publication of the *sizespectra* package [-@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task. Therefore, we recommend using this in all future studies of size spectra relationships.

@@

### Range of body sizes = 1 to 100

![Relationship estimate](figures/PLB_small_m_relationship_density.png)

@@

SI Table. This table shows published estimates of the variation in size spectra slopes (or exponents) in empirical studies. It is unclear how to directly compare estimates of the slope with different methods.However, the published estimates here range from ~0.1 to 0.2 across the gradients studied. For comparison, the 2.5-95% quantiles around the relationship estimate for the MLE method were ~0.1, whereas for the ELBn and NAS method they were ~0.25 and ~0.2, respectively. b_diff is the change in estimate (b-low - b-min). System refers to stream communities or mesocosm experiments. Method: MLE = maximum likelihod estimate, ELB = equal logarthmic binning, the number before indicates the number of bins used. The normalization process shifts the estimates by an absolute value of 1.0. Hence, direct comparison the relative change in normalized and non-normalized studies should not introduce any bias. The O'Gorman et al. 2017 study used average species size and abundance (Local Size Density Relationship, *sensu* White et al. 2007) as opposed to individual size distribution. These methods are related, but it is unclear how to directly compare estimates from each method.

```{r, echo=FALSE}
df <- read.csv("lit comparison.csv")
knitr::kable(df,
             col.names = c("Citation",
                           "$\\beta_1$ Difference",
                           "Error", 
                           "Error Type",
                           "$\\beta_{low}$",
                           "$\\beta_{high}$",
                           "System",
                           "Driver", 
                           "Method"))

```{r, echo=TRUE}
# kable() above looks awful
# here is another bad option
# if we decide to include this I'll make it look nice in excel. 
read.csv("lit comparison.csv")
```

>>>>

---
title: "Manuscript: Detecting differences in Size Spectra"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: 
  - Justin Pomeranz:
      email: jfpomeranz@gmail.com
      institute: cmu
      correspondence: true
  - name: James R. Junker
    institute: [glrc, lumcon]
  - name: Vojsava Gjoni
    institute: usd
  - name: Jeff S. Wesner
    institute: usd
institute:
  - cmu: Colorado Mesa University, Grand Junction, CO, USA
  - glrc: Great Lakes Research Center, Michigan Technological University, Houghton, MI USA
  - lumcon: Louisiana Universities Marine Consortium, Chauvin, LA USA
  - usd: Dept. of Biology, University of South Dakota, Vermillion, SD, USA
output:
   pdf_document:
     toc: yes
     keep_tex: true
     pandoc_args:
     - '--lua-filter=scholarly-metadata.lua'
     - '--lua-filter=author-info-blocks.lua'
   word_document:
      reference_docx: working_docx_template.docx
      toc: no
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
header-includes:
  - \usepackage{lineno}
  - \usepackage{amsmath}
  - \usepackage{indentfirst}
  - \linenumbers
  - \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
indent: true
linestretch: 1
bibliography: detecting-size-spectra-difference.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

@@

@@

Body size distributions are a fundamental characteristic of communities. In general, size-abundance relationships are expected to scale with the value of -0.75 across all biological communities [@Damuth1981; @Damuth1991; @Damuth1998]. This is thought to be a consequence of simple size-dependent metabolic constraints on organisms’ energy use predicted by the metabolic theory of ecology [@Nee1991; @Brown2004]. The remarkable consistency of these relationships across spatiotemporal scales and ecosystems has led them to be recommended as a "universal" indicator of ecological status [@Petchey2010]. Variation in size-abundance relationships have been documented through space [@pomeranz2022], time [@evans2022], seasonality [@mcgarvey2018], and in response to human activities [@jennings2004]. Likewise, variation in size spectra relationships have been used to explain fundamental differences in how communities are organized. For example, external resource subsidies "bend the rules" and allow higher abundances of large body sizes than would be expected [@perkins2018]. However, recent research has shown that these results may be an artifact of how the data were treated. Edwards et al. (2020) reanalyzed a time series of marine fisheries data and found that previously reported changes through time were actually dependent on the methodology used.

Individual size distributions (ISD *sensu* [@White2007], also referred to as abundance size spectra) are one of the size-abundance relationships commonly used. ISDs represent frequency distributions of individual body sizes within a community, regardless of taxonomy. Generally, there is a negative relationship between individual body size (M) on the *x*-axis  and abundance (N) on the *y*-axis. Theoretical and empirical data support this relationship being described as a simple power law with exponent $\lambda$ in the form of $N \sim M^{\lambda}$, where $\lambda = 2$ [@Sheldon1972; @Andersen2006a]. Commonly, $N$ is the count of body sizes grouped into bins, and $\lambda$ is estimated as the slope from OLS regressions in log-log space of $N$ and the mid point of the body size bin $M_{bin}$: $\log(N_{count}) = \lambda \log(M_{bin})$. Myriad binning methods have been proposed, including linear and logarithmic bin widths. Likewise, some methods rely on the absolute counts in the bins and others employ normalization techniques such as dividing the count by the bin width (especially common with logarithmic binning). As an alternative to binning methods, $\lambda$ can be estimated directly on un-binned data using maximum likelihood techniques.

Previous work has shown that the estimates of $\lambda$ differ between MLE and size-binned OLS techniques. Size-binned OLS methods are particularly sensitive to decisions made in the binning process. Simulation studies have shown that MLE offers consistently more accurate estimates of $\lambda$ [@white2008; @edwards2017], and reanalysis of empirical data also indicates that the conclusions are dependent on the methodology used [@white2008; @edwards2020]. However, recent empirical analysis of stream macroinvertebrate communities across the National Ecological Observatory Network (NEON, USA) showed that while the estimates of $\lambda$ varied based on method, the relative change across the environmental gradient was consistent across methods [@pomeranz2022]. While there is a growing consensus that MLE methods offer more reliable estimates of $\lambda$, and binning methods result in biased estimates, it remains unclear if these biases are consistent and systematic or stochastic, and whether or not the relative change in ISD parameters is consistent across space and time. In other words, if the data within a study are all treated the same, does a relative change of size-binned OLS slope parameters of 0.1 coincide with a relative change of MLE $\lambda$ estimates of 0.1?

@@

For the first binning method, we created 6 equal logarithmic bins covering the range of body sizes. This method has been used extensively in previous studies. For example see [@perkins2018; @martinez2016; @dossena2012]. In the present study, the count in each bin was normalized by dividing by the bin width to account for the unequal bin sizes. The process of normalizing shifts the estimate by -1. In other words, an un-normalized estimate of -0.75 would result in an estimate of -1.75 when normalizing the data [@sprules2015; @edwards2017; @pomeranz2022]. Throughout the manuscript, the normalized equal logarithmic binning method will be referred to as ELBn.

### Log~2~ Bins: NAS
The second method was similar to ELBn, but bins of log~2~ widths are used. The number of bins is dependent on the range of body sizes present in the data. When working with empirical data with different size ranges this can alter the number of bins per site, and it is recommended to construct the bins based on the global min and max body size present. However, since the data here were simulated from a known size range, the number of bins for each site is identical. The count in each bin is normalized in the same way as described above. This method has been  used extensively in the literature e.g., [@jennings2004; @Jennings2002; @sprules2015; @pomeranz2019; @mcgarvey2018] and is referred to as the Normalized Abundance Spectrum (NAS).

@@

There was considerable variation in the $\lambda$ estimate across methodologies (Figure 1). The distribution of estimates from the MLE method was always symmetrical and centered at the known value of $\lambda$ (Figure 1). The distribution of estimates from the binning methods were generally wider and occasionally asymmetrical or bimodal. On average, the 95% confidence intervals produced by the NAS and ELBn methods were ~2 times wider than those produced by MLE (Table 1), indicating greater consistency of estimates from MLE. Similarly, estimates of  $\lambda$ deviated from the true $\lambda$ by an average of 0.04 or 0.05 absolute units for the NAS and ELBn methods, more than 4 times higher than deviation (0.01) for the MLE (Table 1).

@@

```{r, echo=FALSE}
t1 <- read.csv("tables/bias_table.csv")
knitr::kable(t1,
             col.names = c("Target",
                           "Method",
                           "n",
                           "Median Range of 95% CI",
                           "Median Absolute Deviation",
                           "SD Absolute Deviation"),
                           caption = "Table 1. Summary of three methods in recapturing the known values of lambda or the regression slopes simulated in this study. Performance is determined by comparinged by uncertainty (range of 95% CI's) and the absolute distance of the model estimates from the known estimates (median and sd of the difference). Values are summarized across all n = 9000 or 6000 simulated datasets. See figures for more specific comparisons.")

@@

The MLE method (Figure 3, blue) recaptured the known slope value in each of the “windows.”, with median absolute deviation of ~0.01 units (Table 1). By contrast, the binning methods systematically overestimated the known slope (Figure 3), with median absolute deviations >4 times higher than the MLE. Similarly, uncertainty in the slope estimates derived from binning methods was twice that of the uncertainty in the MLE method (Table 1).

@@

All methods recaptured the correct sign of the slopes, yielding qualitative consistency (Figure 4). However, the binning methods again systematically underestimated the true value of the slope by ~0.05 units (Figure 5). Likewise, uncertainty in the slope estimates was always greater in the binning methods, with the width of the distributions increases with stronger relationships across a hypothetical gradient. By comparison, the MLE showed no evidence of bias and was always centered at the known value with relatively narrow variation.

![Figure 4. Individual regressions for each rep (N = 1000) and method (columns) for three different known relationship values (rows from bottom to top): 0, 0.25, and 0.5](figures/vary_beta_plot.png)

@@

Both empirical data sets yielded similar results to the simulated data; the direction of the coefficients (i.e., $\beta_1$ coefficients) were the same among methods, with positive slopes across the AMD gradient (Fig. 6A) and negative slopes across the temperature gradient. However, the magnitude of change differed between methods. The AMD slopes ranged from ~0.62 with MLE to ~0.78 with NAS (Figure 6b), while temperature slopes ranged from -0.0058 with MLE to -0.0019 with ELBn (Figure 6d). As with simulated data, slope uncertainty ($\pm$ 1 SD) was larger in the binning methods, particularly the ELBn method (Fig. 6B). Likewise, the size spectra parameters consistently increase (become steeper) with increasing temperature across the NEON sites (Fig. 6C).

@@

@@

The relationship between body size and abundance has been extensively studied in a wide range of taxa inhabiting both terrestrial and aquatic ecosystems (reviewed by [@brown1995; GastonBlackburn2000; @White2007]). Empirical data shows generally consistent patterns and can be explained by the metabolic theory of ecology [@Brown2004].  Measuring parameters describing the decline in abundance with increasing body size in communities is being done with increasing frequency across ecology. Previous work has investigated the accuracy and inherent biases associated with different estimation methods [@White2007; @edwards2017; @edwards2020]. However, how these inaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detect variation in size spectra parameters across environmental gradients with confidence. The most important outcome of our work is that binning methods not only generate biased $\lambda$ values for individual data sets, but that bias carries over to affect the parameters of subsequent regressions that use those $\lambda$’s as response variables. This makes it challenging to understand how $\lambda$ varies in response to environmental gradients if binning is used to estimate size spectra exponents.

Binning methods are easy to use and interpret, which most likely accounts for their wide use in ecological studies. However, aggregating individuals into logarithmic bins removes a large source of the variation within the data by collapsing body size variation into a single value within each bin. For example, all individuals placed into a bin that ranges from 2-4 grams of mass are all treated as having a mass of 3 grams, the midpoint of that bin. Likewise, a single abundance value is taken for each bin, despite that fact that there is almost certainly variation in the abundance of individuals that weigh \~2, \~3, or \~4 grams. Moreover, the number of bins that can be produced by any dataset is limited most often to less than six. That means a subsequent log-log regression used to estimate $\lambda$ would contains n = 6 data points, even though there likely 100’s or 1000’s of individual body sizes available in a typical body size data set. By homogenizing the data in this way, binning methods produce noisier results than MLE, since binning is akin to deleting information that the model could otherwise use. By contrast, the MLE uses all the individual body size data to directly estimate $\lambda$, meaning that it not only produces more accurate estimates, but does so with less certainty than binning, even when the underlying data sets are identical.

The error in estimating individual $\lambda$’s was compounded when estimating linear models with those $\lambda$’s as response variables. The MLE method also resulted in the most accurate $\beta_{env}$ estimates. Furthermore, variation around the estimate was always smallest with the MLE method.

Although there were differences in magnitude of the empirical relationship coefficients, they were generally in the correct direction and of a similar magnitude. This suggests that previously reported significant changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible. Given that all of the data within a study is treated identically, the over all change in size spectra parameters is likely reasonable. However, the biases and inconsistencies in relationship estimates presented here suggest that it would be difficult if not impossible to directly compare the relative changes across different published studies which use different methods.

The publication of individual body size data with future studies of size spectra would greatly aid in our ability to generalize changes to this fundamental aspect of community organization across spatiotemporal scales and in response to environmental conditions.

@@

The MLE method outperformed both of the binning methods under nearly any measure. With the publication of the *sizespectra* package [-@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task. Therefore, we recommend using it in all future studies of size spectra relationships rather than binning.

@@

@@

SI Table. This table shows published estimates of the variation in size spectra slopes (or exponents) in empirical studies. It is unclear how to directly compare estimates of the slope with different methods. However, the published estimates here range from ~0.1 to 0.2 across the gradients studied. For comparison, the 2.5-95% quantiles around the relationship estimate for the MLE method were ~0.1, whereas for the ELBn and NAS method they were ~0.25 and ~0.2, respectively. b_diff is the change in estimate (b-low - b-min). System refers to stream communities or mesocosm experiments. Method: MLE = maximum likelihod estimate, ELB = equal logarthmic binning, the number before indicates the number of bins used. The normalization process shifts the estimates by an absolute value of 1.0. Hence, direct comparison the relative change in normalized and non-normalized studies should not introduce any bias. The O'Gorman et al. 2017 study used average species size and abundance (Local Size Density Relationship, *sensu* White et al. 2007) as opposed to individual size distribution. These methods are related, but it is unclear how to directly compare estimates from each method.

Supplementary table comparing with other published results will be submitted as a separate file for formatting.
